---
title: "Teaching demo: Intro to simulation"
author: "Ciaran Evans"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Context

* This lecture has been used as the first lecture in STA 279 Statistical Computing

\vspace{0.2cm}

* Students in Statistical Computing have taken STA 112 and are familiar with some R fundamentals; no other background is assumed

\vspace{0.2cm}

* First unit of Statistical Computing is on simulation studies
    * Allows review of some R basics
    * Provides context and motivation for fundamental computing concepts: data types, iteration, good coding practices
    
## Today's lesson

* **Learning goal:** by the end of this lesson, students will be able to implement a short simulation to answer a probability question

\vspace{0.15cm}

* Topics reviewed or introduced:
    * Planning simulations
    * Vectors in R
    * Iteration
    
\vspace{0.15cm}

* Student participation:
    * Short neighbor/group discussions
    * Dialogue and questions throughout
    * Your turn: activity at the end of the lesson
    

## Warm-up question

**Problem:** 10 people are at a party, and all of them are wearing hats. They each place their hat in a pile; when they leave, they choose a hat at random. What is the probability at least one person selected the correct hat?


\vspace{0.5cm}

**Question:** Work with your neighbor to discuss the following question:

-   Without calculating probabilities, how could you design an experiment to estimate this probability?

## Designing an experiment

## Step 1: representing the hats

```{r, echo=T}
hats <- 1:10

hats
hats[3]
```

-   `hats` is a **vector**, containing the numbers 1 to 10
-   entries in a vector are accessed by their index

## Step 2: everyone draws a random hat

```{r, echo=F}
set.seed(7)
```


```{r, echo=T}
hats <- 1:10
randomized_hats <- sample(hats, size = 10, 
                          replace = FALSE)

hats
randomized_hats
```

-   The `sample` function creates a random sample from a vector
-   How many people selected their original hat?

## Step 3: check who got their original hat

\small

```{r, echo=T}
hats
randomized_hats
```

\pause

```{r, echo=T, eval=F}
hats == randomized_hats
```

\begin{verbatim}
FALSE FALSE FALSE  TRUE FALSE  TRUE FALSE FALSE FALSE FALSE
\end{verbatim}

\pause

```{r, echo=T}
# TRUE is 1, FALSE is 0
sum(hats == randomized_hats)
```

\pause

```{r, echo=T}
# did at least one person get their hat?
sum(hats == randomized_hats) > 0
```


## Code so far

```{r, echo=T, eval=F}
hats <- 1:10
randomized_hats <- sample(hats, size = 10, 
                          replace = FALSE)
```

```{r, echo=T}
sum(hats == randomized_hats) > 0
```

- In this case, at least one person received their original hat!
- Is this a good estimate of the *probability*?

## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T, eval=F}
nsim <- 10000 # number of simulations 
for(i in 1:nsim){
  
  
}
```

## Loop example

```{r}
for(i in 1:5){
  print(3)
}
```

## Loop example

```{r, eval=F}
for(i in 1:5){
  print(i)
}
```

## Loop example

```{r}
for(i in 1:5){
  print(i)
}
```

## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T}
nsim <- 10000 # number of simulations
hats <- 1:10

for(i in 1:nsim){
  randomized_hats <- sample(hats, size = 10, 
                            replace = FALSE)
  print(sum(hats == randomized_hats) > 0)
}
```


## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T}
nsim <- 10000 # number of simulations
hats <- 1:10
results <- rep(NA, nsim) # vector to store results

for(i in 1:nsim){
  randomized_hats <- sample(hats, size = 10, 
                            replace = FALSE)
  results[i] <- sum(hats == randomized_hats) > 0
}

head(results)
```

## Step 4: iteration

A `for` loop repeats code many times:

```{r, echo=T}
nsim <- 10000 # number of simulations
hats <- 1:10
results <- rep(NA, nsim) # vector to store results

for(i in 1:nsim){
  randomized_hats <- sample(hats, size = 10, 
                            replace = FALSE)
  results[i] <- sum(hats == randomized_hats) > 0
}

mean(results)
```


## Class activity

For the remainder of class, work with a neighbor on the class activity (link below and on the course website):

[https://sta279-example.github.io/class_activities/ca_lecture_1.html](https://sta279-example.github.io/class_activities/ca_lecture_1.html)


## What comes next?

* Continuing probability simulations (gambler's ruin, airplane seating, Monty Hall problem, etc.)
    * setting seeds
    * good coding practices
    * `for` and `while` loops
    * nested loops
    * `if...else if...else` statements
    
\vspace{0.2cm}    

* Statistical simulations
    * answering questions about linear regression models (e.g., does constant variance matter?)
    * ADEMP framework^[["Using simulation studies to evaluate statistical methods"](https://onlinelibrary.wiley.com/doi/10.1002/sim.8086) (Morris *et al.* 2019)]
    * introduction to writing functions


## What is the exact probability?
